Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex

Neurons communicate through short spikes, characterized by a high
frequency, while low frequencies of the signal correspond to the slow
decay of fluorescence. To have a signal that only has high magnitude around instants where the spikes occur, the second step of our pipeline transforms the time series into its backward
difference
\begin{align}
g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i \label{eq:high-pass-filter},
\end{align}
shown in the plot of Figure \ref{fig:hp_curve}:

@@

The objective of the last step of our filtering procedure is to decrease the importance of spikes that occur when there is a high global activity in the network with respect to  spikes that occur during normal activity. Indeed, we have conjectured that when a large part of the network is firing, the rate at which observations are made is not high enough for being able to detect interactions and, that it would therefore be preferrable to lower their importance by changing appropriatly their magnitude. Additionnally, it is well-known that neurons may also spike because of a global high activity (see \cite{stetter2012model}). In such  context, detecting pairwise neuron interactions from the firing activity is meaningless. 
Accordingly, we have chosen to apply to the signal outputted by $h$ the following function  
\begin{align}
 w(x^{t}_i) &= (x^{t}_i + 1 )^{1 + \frac{1}{\sum_{j} x^{t}_j}}.
\end{align}
which effect is to magnify the importance of spikes that occur in case of low global activity (measured by $\sum_{j} x^{t}_j$).

>>>>

Furthermore, it can be observed that neurons communicate through short spikes, characterized by a high
frequency, while low frequencies of the signal mainly correspond to the slow
decay of fluorescence. To have a signal that only has high magnitude around instances where the spikes occur, the second step of our pipeline transforms the time series into its backward
difference
\begin{align}
g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i \label{eq:high-pass-filter},
\end{align}
shown in the plot of Figure \ref{fig:hp_curve}:

@@

The objective of the last step of our filtering procedure is to decrease the importance of spikes that occur when there is a high global activity in the network with respect to  spikes that occur during normal activity. Indeed, we have conjectured that when a large part of the network is firing, the rate at which observations are made is not high enough for being able to detect interactions and, that it would therefore be preferrable to lower their importance by changing appropriatly their magnitude. Additionnally, it is well-known that neurons may also spike because of a global high activity (see \cite{stetter2012model}). In such  context, detecting pairwise neuron interactions from the firing activity is meaningless. 
Accordingly, we have chosen to apply to the signal outputted by $h$ the following function  
\begin{align}
 w(x^{t}_i) &= (x^{t}_i + 1 )^{1 + \frac{1}{\sum_{j} x^{t}_j}}.
\end{align}
which effect is to magnify the importance of spikes that occur in case of low global activity. Note that the particular case where $\sum_{j} x^{t}_j$ equals $0$ is solved by setting $w(x^{t}_i) = 1$.
